### 2020-08-12 paper/table/graphs.tex
evaluation, finally

\begin{tabular}{lrrr}
\toprule
 &          Nodes &          Edges & Preprocessing \\ & $[\cdot 10^6]$ & $[\cdot 10^6]$ &           [s] \\
\midrule
OSM Ger &       16.2 &       35.4 &                         295.2 \\
TDEur17 &       25.8 &       55.5 &                         292.6 \\
TDGer06 &        4.7 &       10.8 &                          58.9 \\
\bottomrule
\end{tabular}