### 2021-11-16 paper/ch_potentials.tex
grammar checking

% \documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
\documentclass[manuscript,review]{acmart}

\usepackage[utf8]{inputenc}

% \usepackage{amssymb}
% \usepackage{amsmath}
\newcommand{\cmark}{$\bullet$}
\newcommand{\xmark}{$\circ$}

\usepackage{todonotes}
% \usepackage{booktabs}
\usepackage{multirow}
\usepackage[algo2e,vlined]{algorithm2e}
\usepackage{todonotes}

\usepackage{tikz}
\usetikzlibrary{automata,arrows,positioning,calc}
\tikzstyle{node}=[circle,inner sep=0.5mm,minimum size=6.5mm,draw = black]

\newcommand*{\dist}{\mathcal{D}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

\begin{document}

\title{Using Incremental Many-to-One Queries to Build a Fast and Tight Heuristic for A* in Road Networks}

\author{Ben Strasser}
\email{academia@ben-strasser.net}
\affiliation{}

\author{Tim Zeitz}
\email{tim.zeitz@kit.edu}
\orcid{0000-0003-4746-3582}
\affiliation{%
  \institution{Institute of Theoretical Informatics, Algorithmics I, Karlsruhe Institute of Technology}
  \city{Karlsruhe}
  \country{Germany}
}

% \author{Ben Strasser}{Germany}{academia@ben-strasser.net}{}{}
% \author{Tim Zeitz}{Karlsruhe Institute of Technology, Germany}{tim.zeitz@kit.edu}{https://orcid.org/0000-0003-4746-3582}{}

% \authorrunning{B. Strasser and T. Zeitz}
\renewcommand{\shortauthors}{B. Strasser and T. Zeitz}
% \Copyright{Ben Strasser and Tim Zeitz}

\begin{abstract}
We study exact, efficient and practical algorithms for route planning applications in large road networks.
Routing applications often require integrating the current traffic situation, planning ahead with traffic predictions for the future, respecting forbidden turns, and many other features depending on the exact application.
While Dijkstra's algorithm can be used to solve these problems, it is too slow for many applications.
A* is a classical approach to accelerate Dijkstra's algorithm.
A* can support many extended scenarios without much additional implementation complexity.
However, A*'s performance depends on the availability of a good heuristic that estimates distances.
Computing tight distance estimates is a challenge on its own.
On road networks, shortest paths can also be quickly computed using hierarchical speedup techniques.
They achieve speed and exactness but sacrifice A*'s flexibility.
Extending them to certain practical applications can be hard.
In this paper, we present Lazy RPHAST, an algorithm to efficiently compute many distance estimates towards a single target and use it as an A* heuristic.
As it uses Contraction Hierarchies (CH), a hierarchical technique, we call our A* heuristic CH-Potentials.
Our approach allows decoupling the supported applications from the hierarchical speed-up technique.
Additionally, we describe A* optimizations to accelerate the processing of low degree nodes, which often occur in road networks and present a new pruning criterion for symmetrical bidirectional A*.
An extensive experimental study confirms the practicality of our approach for many applications.
\end{abstract}

\ccsdesc[500]{Theory of computation~Shortest paths}
\ccsdesc[300]{Mathematics of computing~Graph algorithms}
\ccsdesc[500]{Applied computing~Transportation}

\keywords{route planning, shortest paths, realistic road networks}

\maketitle

\section{Introduction}
\label{sec:intro}
The past decade has seen a plethora of research on route planning in large street networks~\cite{bdgmpsww-rptn-16}.
Routing a user through a road network can be formalized as the shortest path problem in weighted graphs.
Nodes represent intersections.
Roads are modeled using edges.
Edges are weighted by their traversal times.
The problem can be solved with Dijkstra's algorithm~\cite{d-ntpcg-59}.
Unfortunately, on continental sized networks, it is too slow for many applications.
Thus, speed-up techniques have been developed.
One popular example are Contraction Hierarchies~(CH)~\cite{gssv-erlrn-12}.
They have been used successfully in many real world applications.
A CH exploits the inherent hierarchy of road networks.
In a preprocessing step, additional shortcut edges are inserted, which allow skipping unimportant parts of the network at query time.
Another popular example is Multi-Level-Dijkstra~(MLD)~\cite{swz-umlgt-02} also known as CRP~\cite{dgpw-crprn-13}.
It is also used in practice~\cite{bingblog}.
MLD also uses shortcut edges.
Both approaches achieve speed-ups of at least three orders of magnitude over Dijkstra's algorithm.

\begin{figure}
\centering
\includegraphics[width=.6\columnwidth]{fig/searchspace_st.png}
\caption{Nodes explored by A*. Color indicates the node removal order from the queue. Blue was removed first. Next is green. Red was removed last.}
\label{img:search-space}
\Description[A map where nodes visited by A* are marked.]{
A map of Karlsruhe where nodes visited by A* are marked.
The shape of the area with the visited nodes is similar to a drop.
The source node roughly lies in the center of the drop, the target node at the tip.
}
\end{figure}

Unfortunately, for many real world applications, this basic graph model is too simplistic.
For realistic routing, many additional features need to be considered.
This includes turn costs and restrictions, live traffic, user preferences, and traffic predictions. %, temporary driving bans for certain types of vehicles.
Some applications may have additional application-specific requirements.
Extending Dijkstra's algorithm to support these features is usually easy.
Extending hierarchical speed-up techniques is often possible, too.
However, the algorithm development is vastly more complex.
For every feature, dedicated research papers exist that extend CH and other approaches.
Supporting the combination of several features is even harder.
For example, we are not aware of any work combining all features mentioned above.
In this article, we describe an algorithmic building block, that allows handling the combination of all above-mentioned features -- and probably more.

Our approach decouples the development of extensions from the hierarchical speed-up technique by utilizing the A* algorithm~\cite{hnr-afbhd-68}.
A* is a goal-directed variant of Dijkstra's algorithm.
See Figure~\ref{img:search-space} for an example of nodes traversed during an A* search.
A* uses a \emph{heuristic} to guide the search towards the goal.
A heuristic is a function that maps a node to an estimate of the remaining distance to the goal.
A*'s running time crucially depends on how tight this estimate is.
Further, evaluating the heuristic must be fast.
% The free flow travel time to the goal is a valid heuristic for most features.
% In many settings, it is the tightest lower bound available during preprocessing, i.e., independent of real-time data and user settings.
% However, to be useful we must be able to evaluate the free flow heuristic \todo{ist free flow heuristic besser als tight heuristic?} quickly and exactly.
In this article, we introduce Lazy RPHAST and build on it CH-Potentials, a fast heuristic with tight estimates.
Internally, the heuristic uses a CH.
Fortunately, this is an implementation detail from the perspective of the A*.
To support a new feature, we only need to modify the A* algorithm.
The heuristic core containing the CH remains untouched.
Extending A* is vastly easier than extending a CH.
This enables us to design algorithms for a multitude of features.

This article is the extended version of our conference paper ``A Fast and Tight Heuristic for A* in Road Networks''~\cite{strasser_et_al:LIPIcs.SEA.2021.6}.
The paper introduced the basic CH-Potentials heuristic, some query optimizations for handling of low-degree nodes, common in road networks, and some applications of CH-Potentials.
In this article, we discuss and evaluate Lazy RPHAST, the central ingredient of CH-Potentials, in greater detail and introduce an improved variant for Customizable Contraction Hierarchies.
% Furthermore, we evaluate Lazy RPHAST in the context of one-to-many and many-to-many problems.
The original paper only discussed unidirectional CH-Potentials.
In this journal version, we also discuss a bidirectional variant and introduce an improved pruning criterion for symmetrical bidirectional A*.
The improved pruning is not specific to CH-Potentials and can also be used with other heuristics.
Compared to the conference paper, the journal version contains a significantly expanded experimental evaluation.

The rest of the article is organized as follows.
In Section~\ref{sec:related_work}, we discuss related works on goal directed search and extensions for realistic applications for hierarchical techniques.
After establishing preliminaries in Section~\ref{sec:preliminaries}, we introduce the incremental many-to-one problem and our algorithm to solve it efficiently in Section~\ref{sec:lazy-rphast}.
CH-Potentials, our new distance estimation function is introduced in Section~\ref{sec:main-algo}.
There, we also discuss our improvements to the handling of low-degree nodes (Section~\ref{sec:low-deg-improvment}) and for symmetrical bidirectional A* (Section~\ref{sec:bidir_astar}).
% Section~\ref{sec:extensions} presents a multitude of applications of CH-Potentials.
In Section~\ref{sec:extensions}, we demonstrate CH-Potential's flexibility, by describing how to apply the approach to different practical applications.
Finally, in Section~\ref{sec:experiments}, we present an experimental evaluation of our approach.

\section{Related Work}\label{sec:related_work}

There is a lot of work that extends hierarchical speed-up techniques to more complex settings~\cite{bdgmpsww-rptn-16}.
For example, in~\cite{gv-errnt-11} turn information is integrated into CH.
The proposed approach achieves fast queries, but preprocessing becomes an order of magnitude slower compared to classical CH.

A considerable amount of research and engineering effort has been put into studying the combination of traffic predictions with CH.
Several papers~\cite{bdsv-tdch-09,bgns-tdcha-10,klsv-dtdch-10,bgsv-mtdtt-13} and an entire dissertation~\cite{b-tdrpc-14} have been published on the subject.
Different variants with trade-offs regarding exactness, query speed and space consumption were proposed~\cite{bgsv-mtdtt-13}.
Recently, a new approach has been published~\cite{swz-sfert-20} which simultaneously achieves competitive results in all three aspects but only at the cost of considerable implementation complexity.

CRP~(Customizable Route Planning)~\cite{dgpw-crprn-13} is an engineered variant of MLD~\cite{swz-umlgt-02} which was developed to allow updating weights without invalidating the entire preprocessing.
For this, a faster, second preprocessing phase is introduced.
It can be run regularly to update weights.
In theory, this enables the integration of live traffic and user preferences.
In practice, live traffic feed data is imperfect.
Computing ``good'' routes without undesired detours due to artifacts in the data requires additional algorithmic extensions~\cite{dss-tarrn-18}.
CRP also supports turn costs.
Integrating traffic predictions into CRP was studied in~\cite{bdpw-dtdrp-16}.
On continental sized networks, TD-CRP can only compute approximate shortest distances (rather than paths).

In~\cite{dsw-cch-15}, CH is extended to Customizable CH (CCH).
CCH also has a second preprocessing phase where weights can be altered.
Supporting turn costs in CCH was studied in~\cite{bwzz-cchtc-20}.
% Turn integration still causes a slowdown of at least factor three to the second preprocessing phase.

Other extensions studied include electric vehicle routing~\cite{DBLP:journals/algorithmica/BaumDPSWZ20,DBLP:conf/aaai/EisnerFS11} or multi-criteria optimization~\cite{fns-opca-14,gks-rpfof-10}.

% Another feature which has received quite some attention is the computation alternative routes~\cite{bdgs-argrn-11,adgw-arrn-13}.
% Even formally defining what makes a good alternative route is non-trivial~\cite{bdgs-argrn-11}.
% A variety of paradigms exists.
% Competitive approaches require careful adjustment of their respective hierarchical speedup technique for the targeted paradigm.
% In~\cite{krs-eepma-13}, the CRP customization is utilized to repeatedly compute shortest paths on graphs with slightly modified weights.
% In~\cite{k-hdara-13}, partial shortest path trees are extracted from CH search spaces to find via-node candidates for alternative routes.
% In~\cite{barth2019alternative}, a multi-criteria CH is extended to compute alternative routes with regards to multiple edge weight functions.

While these works show that it is possible to extend hierarchical approaches, they also show that it is non-trivial.
Further, in every extension the flexibility available at query time is fairly limited.
Combining these hierarchical extensions is an unsolved problem.

CH-Potentials is not the first work to combine hierarchical approaches and A*~\cite{bdsssw-chgds-10,gkw-blwr-07,bdgwz-sfpcs-19}.
However, previous works mostly focused on accelerating hierarchical approaches further rather than exploiting A*'s flexibility.

ALT~\cite{gh-cspas-05,gw-cppsp-05} and CPD-Heuristics~\cite{DBLP:conf/ijcai/BonoGHS19} are the two techniques with high conceptual similarity to CH-Potentials.
% Just as our approach, ALT is A*-based.
ALT has been combined with shortcuts~\cite{bdsssw-chgds-10} and also extended for dynamic graphs~\cite{dw-lbrdg-07} and time-dependent routing~\cite{ndls-bastd-12,dn-crdtd-12}.
% However, ALT's heuristic is not preprocessing-tight.
%
CPD-Heuristics are a combination of A* and Compressed Path Databases (CPD).
A CPD can quickly compute the first edge of a shortest path between any two nodes.
In~\cite{DBLP:conf/ijcai/BonoGHS19}, SRC~\cite{DBLP:conf/socs/StrasserHB14} is used as CPD.
For every distance estimation, a shortest path to the target is computed, whose length is used as the heuristic value.
Unfortunately, the employed CPD's quadratic preprocessing running time is problematic on large street networks.
%
% MtsCopa~\cite{DBLP:journals/tciaig/BaierBHH15} uses a similar idea to CPD-Heuristics but is only applied in a hunter-prey application.
%
In~\cite{DBLP:conf/ijcai/0002UJAKK18} the weighted graph is embedded into Euclidean space using FastMap. % such that distances in space and distances in the graph roughly correspond.
The Euclidean distance is then used as a distance estimate for A*.
% This results in a good but not a preprocessing-tight heuristic.

\section{Preliminaries}\label{sec:preliminaries}

We consider directed graphs $G=(V,E)$ with $n=|V|$ nodes and $m=|E|$ edges with weight functions $w : E \to \mathbb{R}$.
Given nodes $s$ and $t$ we want to obtain a $st$-path $P=(s=v_0,\dots,t=v_k)$ of minimum weight $w(P) := \sum_{i} w(v_{i-1}v_i)$.
We denote this shortest path weight as the distance $\dist_w(s,t)$ between $s$ and $t$.
If there is no path, we set $\dist_w(s,t)=\infty$.

\subsection{Dijkstra's Algorithm}

In this section, we recall the central aspects of Dijkstra's classical shortest path algorithm~\cite{d-ntpcg-59} and introduce notation used throughout this paper.
Dijkstra's algorithm maintains an array of tentative distances $D$ and a priority queue $Q$ of nodes ordered by increasing distance from $s$.
Initially, all distances are set to $\infty$.
The start node distance $D[s]$ is set to zero and the queue is initialized with $s$.
In each iteration, the closest remaining node $u$ is popped from the queue and \emph{settled}.
Outgoing edges $uv$ are relaxed, i.e. $D[v]$ is set to $\min(D[v], D[u] + w(uv))$.
If the tentative distance at the head node $D[v]$ was improved, it's added to the queue or its position in the queue is adjusted.
Once $t$ was settled, $D[t]$ contains the shortest path distance between $s$ and $t$.
The search can stop when $t$ is settled.

Dijkstra's algorithm can also be run from the target $t$ on the reversed graph.
We call this a \emph{backward search}.
Running two Dijkstra searches simultaneously, one from $s$ and a backward search from $t$, until the searches meet is called \emph{bidirectional} search.
Typically, the searches are interleaved by alternating settling a node from each direction.
Another common approach is to always settle a node from the direction with the smaller minimum queue key.
Theoretically, any direction interleaving strategy is possible.
When the sum of the minimum keys in both queues $\overrightarrow{k} + \overleftarrow{k}$ is greater than the so far best known tentative total distance $\mu$ the algorithm can terminate.

\subsection{A* Algorithm}\label{sec:a_star}

A* is a goal-directed variant of Dijkstra's algorithm.
% A* uses a \emph{heuristic} to guide the search towards the goal.
It uses a heuristic function $h_t(v)$ that maps a node $v$ onto an estimate of the distance from $v$ to the target $t$.
A* orders the nodes in the priority queue by $D[v] + h_t(v)$ instead of $D[v]$ as Dijkstra's algorithm does it.
This results in nodes closer to the target being explored earlier.
Dijkstra's algorithm is a special case of A* with $h_t(v)=0$.
A* is equivalent to running Dijkstra's algorithm on the graph with \emph{reduced weights}~\cite{hnr-afbhd-68}.
This reduced weight function is defined as $w_{h_t}(uv) := w(uv) - h_t(u) + h_t(v)$.
A heuristic function is called \emph{feasible}, if $w_{h_t}(e) \geq 0$ for all edges $e$.
If the employed heuristic is feasible, A* is guaranteed to have found the optimal distance after settling $t$.
In this article, we limit our discussion to feasible heuristics.
% TODO more explicit details on A* correctness?

\subsection{Contraction Hierarchies (CH)}

\begin{figure}
\centering
\begin{tikzpicture}[]
\tikzset{auto rotate/.style={to path={let \p1=(\tikztostart),\p2=(\tikztotarget),
\n1={atan2(\y2-\y1,\x2-\x1)},\n2={\n1-10},\n3={\n1+190}
in (\tikztostart.\n2) -- (\tikztotarget.\n3) \tikztonodes}}}

\node (s) at (0,0)     [node] {$s$};
\node (2) at (1.5,1.5) [node] {};
\node (3) at (3.7,1)   [node] {};
\node (4) at (5,3)     [node] {$m$};
\node (5) at (5.75,-0.1)  [node] {};
\node (6) at (6.75,1.2)   [node] {};
\node (7) at (7.75,-0.5)    [node] {};
\node (8) at (8.5,2)   [node] {};
\node (t) at (10,0.5)  [node] {$t$};

\draw[->] (-1,-1) -- (-1,3.5) node [midway, above, sloped] { rank };

\draw (s) edge[red, auto rotate] (2);
\draw[red] (2) -- (3);
\draw[red] (3) -- (4);
\draw[red] (4) -- (5);
\draw[red] (5) -- (6);
\draw[red] (6) -- (7);
\draw[red] (7) -- (8);
\draw (8) edge[red, auto rotate] (t);

\draw (2) edge[blue, auto rotate] (s);
\draw[blue, dashed] (2) -- (4);
\draw[blue, dashed] (4) -- (8);
\draw (t) edge[blue, auto rotate] (8);

\draw[dashed] (4) -- (6);
\draw[dashed] (6) -- (8);
\end{tikzpicture}
\caption{
Solid lines are edges in $G$. Dotted lines are shortcuts. Red is shortest $st$-path in $G$. Blue is equaly long up-down $st$-path in $G^+$. $m$ is the mid node.
}
\label{fig:ch}
\Description[Nodes and edges of a shortest path in the original graph and the corresponding path in the augmented graph.]{
Nodes and edges of a shortest path in the original graph and the corresponding path in the augmented graph.
The y-position on the figure indicates the rank of each node.
Where the original shortest path down and then up, a shortcut edge connects the two higher nodes.
The path in the augmented graph goes only up and then down.
}
\end{figure}

A CH is a two phase technique to efficiently compute exact, shortest paths.
For details, we refer to \cite{gssv-erlrn-12,dsw-cch-15}.
In this section, we give an introduction.

\begin{algorithm2e}
\KwData{$B[v]$: tentative distance from $v$ to target $t$}
\KwData{Minimum priority queue $Q$, also called open list, ordered by tentative distances}
$B[v] \gets +\infty$ for all $v\neq t$;
$B[t] \gets 0$\;
Make $Q$ only contain $t$ with weight $0$\;
\While{$Q$ not empty}{
  $u\gets$ pop minimum element from $Q$\;
  \For{all down-edges $vu$ in $G^+_\ell$}{
    \If{$B[v] > w_\ell(vu) + B[u]$}{
      $B[v]\gets w_\ell(vu) + B[u]$\;
                        Add $v$ or decrease $v$'s key in $Q$ to $B[v]$\;
    }
  }
}
\caption{CH backward search}
\label{algo:ch-backward}
\end{algorithm2e}

In a CH, each node gets a unique rank representing its ``importance''.
By convention, important nodes are usually depicted at the ``top'' while unimportant nodes are positioned at the ``bottom''.
The intuition is that dead-ends are unimportant while highways or bridges are important.
An edge goes \emph{up} when its tail node has lower rank than its head node.
\emph{Down} edges are defined analogously.
An \emph{up-down path} is a path where only one node $m$ is more important than both its neighbors.
$m$ is called the \emph{mid} node.
An \emph{up path} is a path where the last node is the mid node.
Similarly, the first node is the mid node of a \emph{down path}.
% Every up and down path is an up-down path.
%
In the preprocessing phase, first, a node ordering is computed heuristically.
Then, \emph{shortcut} edges are added to the input graph $G$ to obtain $G^+$.
This is done by repeatedly contracting the least important node and adding shortcuts between its neighbors.
%See~\cite{gssv-erlrn-12} for the details.
After the preprocessing, for every pair of nodes $s$ and $t$ there exists a shortest up-down $st$-path in $G^+$ with the same length as a shortest path in $G$.
See Figure~\ref{fig:ch} for a proof sketch.
From every shortest path (red) in $G$, an up-down path of equal length in $G^+$ (blue) can be constructed.
Thus, we can restrict our search to up-down paths in $G^+$.
The search is bidirectional.
The forward search starts from $s$ and only follows up-edges.
Similarly, the backward search starts at $t$ and only follows down-edges in reversed direction.
The two searches meet at the mid node.
Pseudocode for the backward search, i.e., the path from $m$ to $t$, is presented in Algorithm~\ref{algo:ch-backward}.
The forward search works analogously.
%
A CH query is fast, if the number of nodes reachable via only up- or down-nodes is small.
On road networks, this is the case~\cite{gssv-erlrn-12,dgpw-crprn-13}.
On graphs with low treewidth, this is also the case~\cite{dsw-cch-15,hs-gbpo-18}.

% Using the CH query algorithm, we can already give a simple heuristic.
% The heuristic evaluation $h_t(x)$ performs a CH-query from $x$ to $t$.
% This yields tight estimates but a high overhead for the heuristic evaluation.
% While a single CH query is fast, answering one for every node explored in the $A^*$ search is slow.
% Fortunately, we can do better.
% For this we need another component called PHAST~\cite{dgnw-phast-13}.

\subsection{Customizable Contraction Hierarchies (CCH)}
\label{sec:cch-intro}

Some hierarchical techniques were extended to support a three-phase setup.
The phases consist of a slow preprocessing phase, a faster \emph{customization} phase, and fast queries.
The preprocessing phase typically takes hours, the customization seconds, and the query milliseconds.
The preprocessing phase is independent of the weights and only exploits the graph topology.
Weights are introduces in the customization phase.
Paths are computed in the query phase.
This setup is useful to allow for weight changes that are common to all vehicles such as life traffic updates.
The first technique to follow this setup is CRP~\cite{dgpw-crprn-13}.

Customizable Contraction Hierarchies~\cite{dsw-cch-15} (CCH) are a variant of Contraction Hierarchies that uses this three-phase setup.
The key idea is to use a node importance order that is independent of the weight function.
A CCH is also a CH.
The final result of the preprocessing is also an augmented graph $G^+$.
It fulfills all necessary properties for CH query algorithms.
The CH query algorithms can therefore be applied without modifications.
However, the CCH preprocessing grants some additional useful properties.

During preprocessing, an \emph{elimination tree} can be constructed.
The root of the elimination tree is the most important node.
For every other node its parent is its least important upward neighbor.
In~\cite{bcrw-s-16}, it was proven that set of ancestors of a node in the elimination tree is equal to the nodes reachable in a CH search.
This makes it possible to explore the CCH search space with a more efficient shortest path algorithm:
For the forward search, start at the source and follow the elimination tree upward until the root is reached.
All outgoing edges of encountered nodes are relaxed.
The backward search works analogously but starts at the target.
Because this elimination tree based algorithm does not require a queue, it is faster than a Dijkstra-based CCH query.
However, only weight-independent node orderings admit an elimination tree of low height.
The elimination tree query can therefore not be used for a general CH.

\subsection{PHAST}

\begin{algorithm2e}
\KwData{$P[v]$: tentative distance from $v$ to $t$}
Execute Algorithm~\ref{algo:ch-backward}\;
\For{all CH levels $L$ from most to least important}{
	\For{all up-edges $uv$ in $G^+_\ell$ with $u$ in $L$}{
		\If{$P[u] < P[v] + w_\ell(uv)$}{
			$P[u] \leftarrow P[v] + w_\ell(uv)$\;
		}
	}
}
\caption{PHAST basic all-to-one search}
\label{algo:phast}
\end{algorithm2e}

PHAST~\cite{dgnw-phast-13} is a CH extension that computes distances from all nodes to one target node (or vice versa, the reverse case works analogously).
This is sometimes denoted as a \emph{all-to-one} (or \emph{one-to-all}) problem.
The preprocessing phase remains the same as for CH.
The query phase is split into two steps.
The first step is analogue to the CH query:
From $t$, all reachable nodes via reversed down-edges are explored.
Algorithm~\ref{algo:ch-backward} shows this first step.
For the second step, PHAST organizes nodes into \emph{levels}.
These levels correspond to the importance ordering but allow multiple nodes in the same level.
No edge must connect two nodes within the same level.
PHAST then iterates over all CH levels from top to bottom.
In each iteration, all up-edges starting within the current level are followed in reverse.
After all levels are processed, the shortest distances from all nodes to $t$ were computed.
Pseudocode is provided in Algorithm~\ref{algo:phast}.
PHAST is faster than Dijkstra's algorithm on road graphs because it is a better fit for modern processor architectures.
It can also easily be parallelized.
% However, we will not consider parallelization in this paper.
We refer to~\cite{dgnw-phast-13} for an in-depth experimental performance analysis.
% Using PHAST, we can also compute a tight A* heuristic.
% In the query phase, we first run PHAST to compute the distances from every node to $t$ with respect to $w_\ell$ and store the result in an array $H$.
% Next, we run $A^*$ and implement the heuristic as a lookup in the array $H$.

% This PHAST-based algorithm works.
% The $H$ lookup and by extension the $A^*$ search is indeed fast.
% However, the PHAST step before the search is comparatively expensive.
% The reason is that the distances towards $t$ are computed for \emph{all} nodes.
% Ideally, we only want to compute the distances from the nodes explored in the $A^*$ search.

\subsection{RPHAST}

RPHAST~\cite{delling_et_al:OASIcs:2011:3266}, short for Restricted PHAST, is a PHAST extension for efficiently computing distances from a smaller set of source nodes to one target node (again, the reversed case works analogously), solving the \emph{many-to-one} problem.
Given a set of source nodes $S$, the first step is to copy all up-edges reachable only through up-edges from the source nodes into a new graph.
This step is called \emph{selection}.
Then, in the query step, a target $t$ is given and the PHAST algorithm is applied but the downward sweep (second step of the PHAST algorithm) is performed only on the restricted subgraph.
RPHAST is particularly effective when many targets are queried for the same source set $S$.
If the source set changes often, selection times may become problematic.
% RPHAST can also be applied to the \emph{many-to-many} problem, computing a distance matrix between sets $S$ and $T$ of nodes.
% Implemented with SIMD instructions it is, to the best of our knowledge, the fastest currently known algorithm for this problem setting.

\section{The Incremental Many-to-One Problem}\label{sec:lazy-rphast}

In this section, we discuss another very specific problem setting which naturally arises from A* heuristics.
Here, the target node is known in the selection step, but the source nodes are only queried one after another.
We denote this problem as the \emph{incremental many-to-one} problem.
The first step is the \emph{target selection} where the target node is provided.
Then, an arbitrary number of source nodes $s_i$ are given one after another.
For each source the distance to the target has to be computed before the next source will be provided.

We consider the combined running time of the target selection and each incremental query as the total running time to answer an incremental many-to-one query.
Since the target selection time is included, computing the distances to all nodes with Dijkstra or PHAST is too slow.
Also, since the source set is only provided incrementally, RPHAST in its basic form is not well suited to our problem.
Fortunately, we can do better.

\subsection{Lazy RPHAST for Contraction Hierarchies}

\begin{algorithm2e}
\KwData{$B[v]$: tentative distance from $v$ to $t$ as computed by Algorithm~\ref{algo:ch-backward}}
\KwData{$D[v]$: memoized distance from $v$ to $t$, $\bot$ initially}
\SetKwFunction{Dist}{ComputeAndMemoizeDist}
\SetKwProg{Fn}{Function}{:}{}
\Fn{\Dist{$u$}}{
	\If{$D[u] = \bot$}{
		$D[u]\leftarrow B[u]$\;
		\For{all up-edges $uv$ in $G^+_\ell$}{
                        $D[u]\leftarrow\min\{D[u],w_\ell(uv)+\Dist(v)\}$\;
		}
	}
	\Return{$D[u]$}\;
}
\caption{Lazy RPHAST algorithm}
\label{algo:pot}
\end{algorithm2e}

The core idea of our algorithm is to do the (R)PHAST computation lazily and using memoization as depicted in Algorithm~\ref{algo:pot}.
In the target selection, we first run the backward CH search from $t$ to obtain an array $B$.
$B[v]$ is the minimum down $vt$-path distance or $+\infty$, if there is no such path.
$B$ is computed as shown in Algorithm~\ref{algo:ch-backward}.

To compute the distance from a source $s_i$, we recursively compute for all up-edges $(s_i v)$ the distance $D[v]$.
Next, we compute the minimum distance over all up-down paths that contain at least one up-edge using $d = \min_v\{w_\ell(s_i v) + D[v]\}$.
As not all shortest up-down paths contain an up-edge, we set $D[s_i] = \min \{ B[s_i], d \}$.
This calculation is correct, as it computes the minimum up-down $s_it$-path distance in $G^+_\ell$, which corresponds to the minimum $s_it$-path distance in a CH.

\subsection{Lazy RPHAST for Customizable Contraction Hierarchies}

Algorithm~\ref{algo:pot} can be applied to CCH without any modifications.
However, we can also exploit the elimination tree in the Lazy RPHAST algorithm.
To compute the distance $D[v]$ of a node $v$ the distances of all upward neighbors must be final.
In Algorithm~\ref{algo:pot}, we have to compute all these distances recursively.
However, on CCH it is sufficient to ensure that the distance of the elimination tree parent of $v$ is final.
All upward neighbors of $v$ are by definition included in the search space, thus they are also ancestors in the elimination tree.
If the distance of the parent in the elimination tree is final, the distances of all other upward neighbors must already be final, too.
Thus, it is sufficient to iteratively ensure that the distance of the parent in the elimination tree has been computed.
As soon as a node with final distance is encountered, all ancestors are known to have their final distance.
This allows for the more efficient implementation depicted in Algorithm~\ref{algo:cch_pot}.

\begin{algorithm2e}
\KwData{$B[v]$: tentative distance from $v$ to $t$ as computed by Algorithm~\ref{algo:ch-backward}}
\KwData{$D[v]$: memoized distance from $v$ to $t$, $\bot$ initially}
\KwData{$P[v]$: parent node of $v$ in the elimination tree, parent of root is $\bot$}
\KwData{$S$: stack with nodes to compute distances, empty initially, only used to store intermediate data}
\SetKwFunction{Dist}{ComputeAndMemoizeDist}
\SetKw{Break}{break}
\SetKwProg{Fn}{Function}{:}{}
\Fn{\Dist{$u$}}{
  \tcp{Determine for what nodes, the $D$ needs to be filled}
  $v \gets u$\;
	\While{$D[v] = \bot$}{
    Push $v$ onto $S$\;
		\If{$P[v] = \bot$}{
      \Break\;
		}
    $v \gets P[v]$\;
  }
  \tcp{Fill $D$ for those nodes}
  \While{$S$ not empty}{
    $v\leftarrow$ pop top element from $S$\;
		$D[v]\gets B[v]$\;
		\For{all up-edges $vw$ in $G^+_\ell$}{
                        $D[v]\gets\min\{D[v],w_\ell(vw)+D[w]\}$\;
		}
	}
	\Return{$D[u]$}\;
}
\caption{Elimination tree based Lazy RPHAST algorithm}
\label{algo:cch_pot}
\end{algorithm2e}

% \subsection{Many-to-Many with Lazy RPHAST}

% RPHAST was originally developed to answer many-to-many queries.
% It was not designed to be used in conjunction with A*.
% Investigating how Lazy RPHAST performs in the many-to-many setting is therefore interesting.
% In this setting, a set of source and target nodes are given.
% The objective is to compute the distances between all pairs of source and target nodes.
% In this section, we described how Lazy RPHAST can be used to solve this many-to-many problem.
% Our experiments will show, that the achieved performance is comparable to the original RPHAST algorithm.

% To solve the many-to-many problem, RPHAST uses SIMD vector instructions.
% Let $k$ denote the number of elements in one vector.
% RPHAST computes the distances from all source nodes to $k$ target nodes in one run.
% If the target set contains more than $k$ nodes, then the target set is split into smaller sets of $k$ elements\footnote{If the number of targets is not divisible by $k$, the set is increased by adding additional target nodes.}.
% The algorithm is then run for each of these smaller sets.
% Each RPHAST run performs vectorized variants of the selection and query phases.\todo{Is query phase definiert?}

% We translate this setup to the Lazy RPHAST setting as follows.
% First, a set of $k$ target nodes need to be fixed.
% A vectorized selection phase is then performed for this fixed target set.
% After the selection, the source nodes are iteratively revealed.
% Once a source node is revealed, the algorithm must compute the distances to all $k$ targets.
% Only after the distances are computed, the next source node is revealed.

% The core of Lazy RPHAST are the Algorithms~\ref{algo:ch-backward} and~\ref{algo:pot}.
% We describe vectorized variants of these algorithms.
% We replace the $B$ and $D$ arrays of scalars by arrays of $k$-dimensional vectors of scalars.
% To clarify which entities are SIMD vectors, we denote the vectorized arrays by $\hat{B}$ and $\hat{D}$.
% $\hat{D}[s]$ is an SIMD vector $\hat{v}$ of length $k$.
% The $i$-th element $\hat{v}[i]$ of $\hat{v}$ will contain the distance from $s$ to $i$-th selected target.
% We can also denote this distance using $\hat{D}[s][i]$.
% All operations on SIMD vectors are component-wise.
% Adding a scalar to an SIMD vector consists of adding the scalar to every element of the SIMD vector.

% \todo{Rechtschreibung: down edge vs down-edge und up edge vs up-edge. Geht beides aber bitte konsistent.}

% \begin{algorithm2e}
% \KwIn{$\hat{T}$: SIMD vector of $k$ target nodes}
% \KwData{$\hat{B}[x][i]$: tentative distance from $x$ to target $\hat{T}[i]$}
% \KwData{Minimum priority queue $Q$, also called open list, ordered by node rank}
% Make $Q$ an empty queue\;
% $\hat{B}[x][i] \leftarrow +\infty$ for all $x$ and $i$\;
% \For{all $i\in \{0\ldots k-1\}$}{
%   $\hat{B}[\hat{T}[i]][i] \leftarrow 0$\;
%   Add $\hat{T}[i]$ to $Q$\;
% }
% \While{not $Q$ empty}{
%   $y\leftarrow$ pop minimum element from $Q$\;
%   \For{all down-edges $xy$  in $G^+_\ell$}{
%     $\hat{B}[x] \leftarrow \min\{\hat{B}[x], \hat{B}[y] + w_\ell(xy)\}$\;
%     Add $x$ to $Q$ if not in $Q$\;
%   }
% }
% \caption{Vectorized Lazy RPHAST selection}
% \label{algo:simd-ch-backward}
% \end{algorithm2e}

% Vectorizing Algorithm~\ref{algo:ch-backward} needs some adjustments.
% In the original algorithm, the nodes are ordered by distance.
% In the vectorized variant, distances are SIMD vectors and can therefore not be ordered.
% To resolve this problem, we ordered the nodes in the queue not by distance but by CH rank.
% The CH rank is the position of a node in the contraction order. \todo{wurde rank schon eingeführt? Falls ja, dann kann der Satz weg.}
% All other modifications are straight-forward.
% The vectorized version is presented in Algorithm~\ref{algo:simd-ch-backward}.
% Algorithm~\ref{algo:pot} can be vectorized by just replacing the arrays by their SIMD variants.
% The result is illustrated in Algorithm~\ref{algo:simd-pot}.

% \begin{algorithm2e}
% \KwData{$\hat{B}[x][i]$: tentative distance from $x$ to $\hat{T}[i]$ as computed by Algorithm~\ref{algo:simd-ch-backward}}
% \KwData{$\hat{D}[x][i]$: memoized distance from $x$ to $\hat{T}[i]$, $\bot$ initially}
% \SetKwFunction{VecDist}{SIMDComputeAndMemoizeDist}
% \SetKwProg{Fn}{Function}{:}{}
% \Fn{\VecDist{$x$}}{
%   \If{$\hat{D}[x] = \bot$}{
%     $\hat{D}[x]\leftarrow \hat{B}[x]$\;
%     \For{all up-edges $xy$ in $G^+_\ell$}{
%       $\hat{D}[x]\leftarrow\min\{\hat{D}[x],w_\ell(xy)+\VecDist(y)\}$\;
%     }
%   }
%   \Return{$\hat{D}[x]$}\;
% }
% \caption{Vectorized Lazy RPHAST algorithm}
% \label{algo:simd-pot}
% \end{algorithm2e}

% \todo{Hier beginnt der ältere Text}
% Since SSE RPHAST is the fastest known algorithm for many-to-many computations, investigating the Lazy RPHAST performance in this setting seems worthwhile.
% SSE RPHAST starts by performing the selection phase on the full target set.
% % TODO flip directions?
% Second, a query is performed for each source which yields the distances from this source to all targets (RPHAST).
% This second step can be improved by processing $k$ (typically 16) sources at once using SIMD vector instructions (SSE RPHAST).
% Naively translating this approach to Lazy RPHAST would mean selecting one target after another and for each doing the full lazy exploration from all sources.
% However, this approach would explore the upward search space of all the sources multiple times.
% This work seems redundant.

% Luckily, we can do better by integrating a different One-to-Many CH variant -- the Bucket-CH algorithm~\cite{gssv-erlrn-12} -- to select all targets at once.
% The Bucket CH selection phase works as follows:
% % TODO better structure or local outline
% % TODO besser herleiten, wieso kein 2d array
% For each node $x$, we maintain an initially empty list $B[x]$ of pair $(t_i, \textrm{dist}_{\textrm{up}}(x,t_i))$.
% These pairs are called \emph{buckets}.
% These buckets are populated by running a backward CH search (Algorithm~\ref{algo:ch-backward}) from each $t_i$ and inserting the resulting distances into the buckets.
% Now Algorithm~\ref{algo:pot} (or~ Algorithm~\ref{algo:cch_pot}) can compute distances from a given source to \emph{all} targets at once.
% We only have to change $D[x]$ so that it is an array of distances to each $t_i$ instead of a single scalar value.
% After calling the adjusted \texttt{ComputeAndMemoizeDist} for each source, we know the shortest distances between each pair of source and target node.
% This way, the search space of the sources is explored only once.

% The Bucket-CH selection phase still explores the shared parts of the search space of the targets several times.
% This can be improved with the following modified CH backward search:
% All target nodes are inserted into the queue at once.
% The priority queue is ordered by importance instead of distance.
% Distance labels are buckets instead of scalar values.
% The buckets of each target node are initialized with a single entry for the respective node with distance zero.
% When relaxing an edge, the entries in the current node's bucket are increased by the edge's weight and merged with the bucket entries of the head node of the edge.
% By keeping the buckets sorted, this merging can be efficiently implemented with a coordinated linear sweep over both buckets.
% We evaluate our Lazy RPHAST many-to-many algorithm with this improved Bucket-CH selection in Section~\ref{sec:exp_lazy_rphast}.
% % TODO describe improved^2 Bucket CH

\section{A* with CH-Potentials}\label{sec:main-algo}

In this section, we describe how to use the Lazy RPHAST algorithm to build a fast and tight heuristic for A*.
We also discuss CH-Potentials for the bidirectional A* algorithm and propose an improved pruning criterion for symmetric bidirectional potentials.
Further, we present several low degree node optimization which exploit road network characteristics to reduce the overhead of queue operations and heuristic evaluations.
Finally, we introduce an algorithmic framework to apply A* with CH-Potentials to a variety of practical route planning problems.

The \emph{CH-Potentials} heuristic is a straightforward application of the Lazy RPHAST algorithm.
In the beginning of each query, we perform the target selection, a backward CH search, from the target $t$.
The heuristic function $h_t(v)$ is implemented by a call to the \texttt{ComputeAndMemoizeDist} for node $v$.

This yields a \emph{perfect} heuristic, at least as long as the CH preprocessing and the A* algorithm are performed on the same graph with the same weight function.
This case is of course not particularly interesting.
One could just answer the shortest path query directly with a CH query.
This approach becomes useful when the query runs on a \emph{different but related} graph or weight function than the preprocessing.
We will discuss such applications in Section~\ref{sec:extensions}.
But before that, we discuss two extensions/optimizations to A* which we will utilize.

\subsection{Low Degree A* Improvements}\label{sec:low-deg-improvment}

Preliminary experiments showed, that a significant amount of query running time is spent in heuristic evaluations and queue operations.
We can reduce both by keeping some nodes out of the queue, as the heuristic needs to be evaluated when a node is pushed into the queue.
Avoiding pushing low degree nodes into the queue is the focus of this section.
The techniques discussed here are a lazy variant of the ideas used in TopoCore~\cite{DBLP:conf/gis/DibbeltSW15}.

We modify A* by processing low degree nodes consecutively without pushing them into the queue.
Our algorithm uses the undirected degree $d(v)$ of a node $v$.
Formally, $d(v)$ is the number of nodes $u$ such that $(vu)\in E$ or $(uv)\in E$.

Analogous to A*, our algorithm stores for every node $v$ a tentative distance $D[v]$.
Additionally, it maintains a minimum priority queue.
Diverging from A*, not all nodes need be pushed, but every node has a tentative distance.

\subsubsection{Skip Degree Two Nodes}

Our algorithm differs from A* when removing a node $u$ from the queue.
A* iterates over the outgoing edges $(uv)$ of $u$ and tries to reduce $D[v]$ by relaxing $(uv)$.
If A* succeeds, $v$'s weight in the queue is set to $D[v]+h_t(v)$.
Our algorithm, however, behaves differently, if $d(v)\le 2$.
Our algorithm determines the longest degree two chain of nodes $u, v_1,\ldots, v_k, w$ such that $d(v_i)=2$ and $d(w) > 2$.
If our algorithm succeeds in reducing $D[v_1]$, it does not push $v_1$ into the queue.
Instead, it iteratively tries to reduce all $D[v_i]$.
If it does not reach $w$, then only $D$ is modified, but no queue action is performed.
If $D[w]$ is modified and $d(w)>2$, $w$'s weight in the queue is set to $D[w]+h_t(w)$.

As the target node $t$ might have degree two, our algorithm cannot rely on stopping, when $t$ is removed from the queue.
Instead, our algorithm stops as soon as $D[t]$ is less than the minimum weight in the queue.

\subsubsection{Skip Degree Three Nodes}

We can also skip some degree three nodes.
Denote by $u, v_1,\ldots, v_k, w$ a degree two chain as described in the previous section.
If $d(w) > 3$ or $w$ is in the queue, our algorithm proceeds as in the previous section.
Otherwise, there exist up to two degree chains $w, x_1,\ldots,x_p, y$ and $w,a_1,\ldots,a_q,b$ such that $x_1\neq y_k \neq a_1$.
Our algorithm iteratively tries to reduce all $D[x_i]$ and $D[a_i]$.
If it reaches $b$, $b$'s weight in the queue is set to $D[b]+h_t(b)$.
Analogously, if $y$ is reached, $y$'s weight is set to $D[y]+h_t(y)$.
If $y$ respectively $b$ are not reached, our algorithm does nothing.

\subsubsection{Stay in Largest Biconnected Component}\label{sec:largested-biconnected-component}

A lot of nodes in road networks lead to dead-ends.
Unless the source or target is in this dead-end, it is unnecessary to explore these nodes.

In the preprocessing phase, we compute the subgraph $G_C$, called \emph{core}.
$G_C$ is induced by the largest biconnected component of the undirected graph underlying $G$.
We do this using Tarjan's algorithm \cite{t-dfslg2-72}.
For every node $v$ in the input graph $G$, we store the attachment node $a_v$ to the core.
For nodes in the core, $a_v=v$.
We exploit that all attachment nodes are single node separators and the problem can be decomposed along them.

The query phase is divided into two steps.
In the first step, we apply A* with CH-Potentials to $G_C$ combined with the component that contains $s$.
This can be achieved implicitly by removing edges from $G_C$ into other components during preprocessing.
If $t$ is part of $G_C$ or in the same component as $s$, this A* search finds it.
Otherwise, we find $a_t$.
In that case, we continue by searching a path from $a_t$ towards $t$ restricted to $t$'s biconnected component.
The final result is the concatenation of both paths.

\subsection{Bidirectional A*}\label{sec:bidir_astar}

On road graphs, bidirectional search provides a way to halve the running time of Dijkstra's algorithm.
A bidirectional variant of A* therefore also seems desirable.
However, as shown in~\cite{gh-cspas-05}, the necessary modifications are not straightforward.
We revisit bidirectional A* and propose an alternative approach.
Our experiments show that it is competitive with the solution described by~\cite{gh-cspas-05}.

The straightforward idea is to use two heuristics $\overrightarrow{h}_t(v)$ and $\overleftarrow{h}_s(v)$.
The forward search has its queue ordered by $\overrightarrow{D}[v] + \overrightarrow{h}_t(v)$, where $\overrightarrow{h}_t(v)$ estimates the distance $\dist(v,t)$ from $v$ to $t$.
Similarly, the backward search has its queue ordered by $\overleftarrow{D}[v] + \overleftarrow{h}_s(v)$, where $\overleftarrow{h}_s(v)$ is an estimate of the distance $\dist(s,v)$ from $s$ to $v$.
We denote by $\overrightarrow{k}$ and $\overleftarrow{k}$ the minimum queue keys and by $\mu$ the length of the shortest $st$-path found so far.

The problem with this straightforward approach is that these two potentials induce different reduced graphs (see Section~\ref{sec:a_star}).
Thus, in the equivalent bidirectional Dijkstra search each direction would run on a different graph.
This breaks the usual bidirectional Dijkstra stopping criterion.
To the best of our knowledge, no better stopping criterion exists than running \emph{both} searches until the unidirectional stopping criterion is met for one direction.
The forward search can skip nodes already settled by the backward search and vice versa.
However, this straightforward bidirectional A* still performs more work than a unidirectional A*~\cite{gh-cspas-05}.
Thus, on its own it is not a useful algorithm.
However, it can serve as basis for further algorithmic refinements.
The authors of~\cite{gh-cspas-05} refer to this approach as \emph{symmetric} bidirectional A*.

To obtain a bidirectional stopping criterion, the \emph{average potential} is proposed~\cite{gh-cspas-05}.
It combines a forward and a backward heuristic $\overrightarrow{h}_t$ and $\overleftarrow{h}_s$ into a combined \emph{average heuristic} $h_{\overline{st}}$.
The idea is to use a common reduced graph, whose weights are the average weights of the individual reduced graphs.
Both searches run on the same common reduced graph.
This allows stopping the searches when $\overrightarrow{k} + \overleftarrow{k} \geq \mu$.
Formally, $h_{\overline{st}}(v)$ is defined as $(\overrightarrow{h}_t(v) - \overleftarrow{h}_s(v))/2$.
The forward search uses $h_{\overline{st}}(v)$ as its heuristic.
The backward search uses $-h_{\overline{st}}(v)$.
Unfortunately, average potentials have two downsides.
First, evaluating the average potential requires evaluating both $\overrightarrow{h}_t$ and $\overleftarrow{h}_s$.
Evaluating the average heuristic is therefore slower than evaluating just one of the composing heuristics.
Second, the bidirectional stopping criterion comes at the cost of worse estimates for each direction on its own.
$h_{\overline{st}}(v)$ is a worse estimate for $\dist(v,t)$ than $\overrightarrow{h}_t$.
Similarly, $-h_{\overline{st}}(v)$ is a worse estimate for $\dist(s,v)$ than $\overleftarrow{h}_s$.
The second downside can be partially mitigated through pruning with the composing heuristics.
When the forward search scans an edge $uv$ where $\overrightarrow{D}[u] + w(uv) + \overrightarrow{h}_s(v) > \mu$ holds, i.e. the distance plus the estimate of the remaining distance is already greater than the currently known tentative distance, it is not necessary to push $y$ into the queue.
The pruning rule for the backward search is analogous.

To avoid the downsides of the average potential, we revisit symmetric bidirectional A* and propose a new pruning criterion.
We describe the idea for the forward search.
The pruning rule for the backward search is analogous.
The central idea consists of using information from the backward search to prune edges in the forward search.
We do not use the average heuristic.
Instead of a strong stopping criterion, we use a pruning rule that gets stronger the longer the search runs.
Such a pruning rule will eventually prune all remaining branches and stop the search.
The stopping criteria for each direction remain the same (unidirectional) as before.
However, usually the search stops early because the queues are empty.

Let $uv$ be an edge that we relax in the forward search.
Before pushing $v$ into the queue, we apply the new pruning rule.
If we can prove that every path using $uv$ is at least as long as the shortest known path length $\mu$, then we do not have to push $v$.
We therefore want to obtain a lower bound for $\dist(s,u)+w(uv)+\dist(v,t)$.
As $u$ was settled, $\overrightarrow{D}[u]$ contains the shortest path length $\dist(s,u)$, i.e., $\dist(s,u)=\overrightarrow{D}[u]$.
$w(uv)$ is also exactly known as it's just an edge weight.
It remains to lower bound $\dist(v,t)$.
Nodes are removed from the backward queue ordered by $\overleftarrow{h}(v) + \dist(v,t)$.
If $v$ was not yet removed from the backward queue, we know that $\overleftarrow{h}(v) + \dist(v,t) \geq \overleftarrow{k}$.
This gives us the required lower bound, i.e. $\dist(v,t) \geq \overleftarrow{k}-\overleftarrow{h}(v)$.
Thus, $v$ does not have to be pushed if $\overrightarrow{D}[u]+w(uv)+\overleftarrow{k}-\overleftarrow{h}(v)\ge \mu$.
The node $v$ might still be pushed into the queue when there is another edge $wv$ for which the pruning is not possible.
Checking the pruning rule requires evaluating the backward heuristic.
However, pruning is only possible once the searches have met, i.e. $\mu < \infty$.
Before that, each direction only has to evaluate its own heuristic.
Thus, our pruning improves on both downsides of the average potential.

Unless stated differently, for all bidirectional A* variants, we always alternate between removing a node from the forward and the backward queues.
In our experiments, we also evaluate expanding the search with the smaller minimum queue key.
This sounds sensible, but our experiments show in Table~\ref{tab:bidir_switching} that it is never beneficial in terms of running time.

\subsection{Applications}\label{sec:extensions}

% A* is a flexible algorithm.
Many extended problems can be solved using CH-Potentials.
We start by establishing a common formal framework for the use of CH-Potentials.
Then, we exemplary describe some extended routing problems and how to apply CH-Potentials to them.

\subsubsection{Formal Setup: Inputs, Outputs, and Phases}

We consider a variety different applications, with slightly different problem models.
The goal is always to quickly answer many shortest path queries.
For the purpose of describing our framework, we establish a shared notation:
Input to each query are nodes $s$ and $t$, and a graph $G_q$ with query weights $w_q$.
However, the precise formal inputs of the query and what exactly $w_q$ represents depends on the application.
In the simplest case, $w_q$ will be scalar edge weights.
However, this is not a requirement.
It can be any function that computes a weight for an edge.
This function can also take additional parameters from the state of the search.
For example, in the case of live-traffic, $w_q$ represents scalar edge weights.
However, values of $w_q$ might change between queries.
In the case of traffic predictions, $w_q$ is a function which maps the edge entry time to the traversal time and the query takes an additional departure time parameter.

To enable quick shortest path computations, we consider a two phase setup with an additional off-line preprocessing phase before the on-line query phase.
The input to the preprocessing phase is a graph $G_\ell$ with lower bound weights $w_\ell$ and a node mapping function $\phi$.
$w_\ell(e)$ must be a scalar value for every edge $e$ of $G_\ell$.
We require that $w_q(u v)$ is greater or equal to the shortest distance $\dist_\ell(\phi(u), \phi(v))$ from $\phi(u)$ to $\phi(v)$ in $G_\ell$. % with respect to $w_\ell$.
The output of the preprocessing is auxiliary data that enables an efficient heuristic function $h_t(x)$.
$h_t(x)$ is the exact distance from $\phi(x)$ to $\phi(t)$ in $G_\ell$.
In the applications considered in this paper, $w_\ell$ is always the freeflow travel time.

The query phase uses this heuristic in an A* search between nodes $s$ and $t$ on $G_q$ and $w_q$.
Unless stated otherwise, $G_q$ and $G_\ell$ are the same graph and only $w_q$ changes for the queries.
The exact implementation of this A* search depends on the application.
Our approach only provides the heuristic $h_t$ for the A* search.
In contrast, the preprocessing phase remains the same for all applications.

Our heuristic is always \emph{feasible}, i.e. $w_q(u v) - h_t(u) + h_t(v) \geq 0$ holds for all edges.
By requirement and because of the triangle inequality the following holds:
\[
w_q(u v) - h_t(u) + h_t(v) \geq \dist_\ell(\phi(u), \phi(v)) - \dist_\ell(\phi(u), \phi(t)) + \dist_\ell(\phi(v), \phi(t)) \geq 0
\]
Thus, A* will always determine the correct shortest distances.

\subsubsection{Avoiding Tunnels and/or Highways}
\label{sec:no-tunnel-highway}

Avoiding tunnels and/or highways is a common feature of navigation devices.
Implementing this feature with CH-Potentials is easy.
We set $w_\ell$ to the freeflow travel time.
If an edge is a tunnel and/or a highway, we set $w_q$ to $+\infty$.
Otherwise, $w_q$ is set to the freeflow travel time.

\subsubsection{Forbidden Turns and Turn Costs}
\label{sec:no-turns}

The classical shortest path problem allows to freely change edges at nodes.
However, in the real world, turn restrictions, such as a forbidden left or right turn, exist.
Also, taking a left turn might take longer than going straight.
This can be modeled using turn weights~\cite{gv-errnt-11,dgpw-crprn-13,bwzz-cchtc-20}.
% In this section, we first present the extended problem setting.
% Afterwards, we describe how it can be solved with CH-Potentials.
A \emph{turn weight} $w_t$ maps a pair of incident edges onto the turning time or $+\infty$ for forbidden turns.
For CH-Potentials, we use zero as lower bound for every turn weight in the heuristic.
Thus, the graph $G_\ell$ and weights $w_\ell$ for preprocessing is the unmodified input graph without turn weights.

A path with nodes $v_1, v_2,\ldots v_k$ has the following \emph{turn-aware weight}: \[
w_\ell(v_1 v_2) + \sum_{i=2}^{k-1}  w_t(v_{i-1},v_i,v_{i+1}) + w_\ell(v_i v_{i+1})
\]
The objective is to find a path between two given edges with minimum turn-aware weight.
The first term $w_\ell(v_1, v_2)$ is the same for all paths, as it only depends on the source edge.
It can thus be ignored during optimization.

We solve this problem by constructing a \emph{turn-expanded} graph as $G_q$.
Edges in the input graph $G_\ell$ correspond to \emph{expanded nodes} in $G_q$.
For every pair of incident edges $x y$ and $y z$ in $G_\ell$, there is an \emph{expanded edge} in $G_q$ with expanded weight $w_t(x,y,z) + w_\ell(y z)$.
A sequence of expanded nodes in the expanded graph $G_q$ corresponds to a sequence of edges in the input graph $G_\ell$.
The weight of a path in $G_q$ is equal to the turn-aware weight of the corresponding path in $G_\ell$ minus the irrelevant $w_\ell(v_1,v_2)$ term.
Thus, the turn-aware routing problem can be solved by searching for shortest paths in $G_q$.

In this scenario, preprocessing and query use different graphs $G_\ell$ and $G_q$.
We define the node mapping function $\phi$ as $\phi(x y) = y$.
Obviously, $w_q(xy, yz) = w_t(x,y,z) + w_\ell(y z) \geq \dist_\ell(\phi(x y), \phi(y z))$ and this approach yields a feasible heuristic.
Sadly, the undirected graph underlying $G_q$ is always biconnected, if the input graph is strongly connected.
The optimization described in Section~\ref{sec:largested-biconnected-component} is therefore ineffective.
With this setup, CH-Potentials support turns without requiring turn information in the CH.

% To prove that $h'(x,y)$ is consistent, consider the turn-aware weight of a path $v_1,v_2\ldots v_k$ with $v_1=x$, $v_2=y$, $v_{k-1}=p$, and $v_k=q$.
% We can lower bound the expression as follows:
% \[
% \sum_{i=2}^{k-1} w_\ell(v_i,v_{i+1}) \le \sum_{i=2}^{k-1} \underbrace{w_t(v_{i-1},v_i,v_{i+1})}_{0\le} + \sum_{i=2}^{k-1} \underbrace{w_q(v_i,v_{i+1})}_{w_\ell(v_i,v_{i+1})\le}
% \]
% $\sum_{i=2}^{k-1} w_\ell(v_i,v_{i+1})$ is a $yq$-path length in $G$.
% It is no shorter than the shortest $yq$-path length in $G$, which is equal to $h(y)$.
% The heuristic is therefore a lower bound.
% It is consistent because it is derived from exact shortest paths in $G$ with $w_\ell$.

\subsubsection{Predicted Traffic or Time-Dependent Routing}
\label{sec:predicted-traffic}

The classical shortest path problem assumes that edge weights are scalars.
However, in practice, travel times vary along an edge due to the traffic situation.
% The primary reason is traffic.
Recurring traffic can be predicted by observing the traffic in the past.
It is common \cite{bgsv-mtdtt-13,bdpw-dtdrp-16,swz-sfert-20} to represent these predictions as \emph{travel time functions}.
An edge weight is no longer a scalar value but a function that maps the entry time onto the traversal time.
% Performing this prediction is outside of the scope of this paper.
% It is common to refer to routing with predicted traffic as \emph{time-dependent routing}.
% Again, we first formalize the extended problem setting and then describe our solution with CH-Potentials.

In this setting, the query weight $w_q$ is a function from $E\times \mathbb{R}$ to $\mathbb{R}^+$.
$w_q(e, \tau)$ is the travel time through edge $e$ when entering it at moment $\tau$.
The input to the extended problem consists of a source node $s$ and a target node $t$, as in the classical problem formulation.
Additionally, the input contains a source time $\tau_s$.
A path with edges $e_1,e_2\ldots e_k$ is weighted using $\alpha_k$, which is defined recursively as follows:\[
\begin{split}
\alpha_{1} & = w_q(e_1, \tau_s) \\
\alpha_{k} & = \alpha_{k-1} + w_q(e_1, \alpha_{k-1})
\end{split}
\]
The objective is to find a path to $t$ that minimizes $\alpha_k$.

If all travel time functions fulfill the \emph{FIFO property}, this problem can be solved using a straightforward extension of Dijkstra's algorithm \cite{d-aassp-69}.
The necessary modification to A* is analogous.
Without the FIFO property the problem becomes NP-hard \cite{or-tnp-89}.
% The FIFO property states that it is never beneficial to wait at a node before entering an edge.
The FIFO property states that it is not possible to arrive earlier by departing later.
Formally stated, the following must hold $\forall e\in E,\tau\in \mathbb{R},\delta\in \mathbb{R}^+: w_q(e, \tau) \le w_q(e, \tau+\delta) + \delta$.
Our implementation stores edge travel times using piece-wise linear functions.
The A* search uses the tentative distance $\tau$ at a node $u$ when evaluating the travel time of outgoing edges $u v$.
This strategy is very similar to TD-ALT~\cite{ndls-bastd-12,dw-lbrdg-07}.

For the preprocessing, we set $w_\ell(e) = \min_\tau w_q(e,\tau)$, that is the minimum travel time.
% With this setup, we extended CH-Potentials to support time-dependent routing.
By keeping travel time functions out of the CH, we avoid a lot of algorithmic complications compared to~\cite{bgsv-mtdtt-13,bdpw-dtdrp-16,swz-sfert-20,dn-crdtd-12} which have to create shortcuts of travel time functions.

\subsubsection{Live and Predicted Traffic}
\label{sec:live-predicted-traffic}

Beside predicted traffic, we also consider live traffic.
Live traffic refers to the current traffic situation.
It is important to distinguish between predicted and live traffic.
Live traffic data is more accurate for the current moment than predicted data.
It is possible that it differs significantly from predicted traffic, if unexpected events like accidents happen.
% The predicted travel time along an edge was estimated in the past.
% Accidents are examples of such unexpected events.
However, just using live traffic data is problematic for long routes as traffic changes while driving.
At some point, one wants to switch from live traffic to the predicted traffic.
In this section, we first describe a setup with only live traffic and then combine it with predicted traffic.

To support only live traffic, we set $w_\ell$ to the freeflow travel time.
$w_q$ is set to the travel time accounting for current traffic.
As traffic only increases the travel time along an edge, $w_\ell$ is a valid lower bound for $w_q$.
In a real world application, values from $w_q$ could be updated between queries.
This is all that is necessary to apply CH-Potentials in a live traffic scenario.

To combine live traffic with predicted traffic, we define a modified travel time function $w_q$ that is then used as query weights.
Denote by $w_p(e,\tau)$ the predicted travel time along edge $e$ at moment $\tau$.
Further, $w_c(e)$ is the travel time according to current live traffic.
Finally, we denote by $\tau_{\mathrm{soon}}$ the moment when we switch to predicted traffic.
In our experiments, we set $\tau_{\mathrm{soon}}$ to one hour in the future.
We need to make sure that the modified travel time function fulfills the no-waiting property.
For this reason, we cannot make a hard switch at $\tau_{\mathrm{soon}}$.
Our modified travel time function linearly approaches the predicted travel time. % with a bounded slope.
%
Formally, we set $w_q(e,\tau)$ to $w_c(e)$, if $\tau \leq \tau_{\mathrm{soon}}$.
Otherwise, we check whether $w_p(e,\tau_{\mathrm{soon}}) < w_c(e)$ is true.
If it is the case, we set $w_q(e,\tau)$ to $\max\{w_c(e)+(\tau_{\mathrm{soon}}-\tau), w_p(e,\tau)\}$.
Otherwise, we set $w_q(e,\tau)$ to $\min\{w_c(e)-(\tau_{\mathrm{soon}}-\tau), w_p(e,\tau)\}$.
% We then run the algorithm of Section~\ref{sec:predicted-traffic} with $w_q$ as query weight.
In our implementation, we to not modify the representation of $w_p$ but evaluate the formulas above at each travel time evaluation.
We set $w_\ell$ again to the freeflow travel time.

With this setup, CH-Potentials support a combination of live and predicted traffic.
We did not make any modification, that would hinder a combination with other extensions.
Further adding tunnel and/or highway avoidance or turn-aware routing is simple.
This straight-forward integration of complex routing problems is the strength of the CH-Potentials.

\subsubsection{CCH Potentials}

Live traffic can alternatively be supported using a three-phase setup as described in Section~\ref{sec:cch-intro}.
We do this by exchanging the CH for a CCH.
The preprocessing phase is identical as for a CCH and is independent of $w_\ell$.
The $w_\ell$ is fixed during the customization phase.
It can be changed arbitrarily by rerunning the customization.
The A* algorithm is part of the query phase.
The weights $w_q$ used in the query phase must not be smaller than the $w_\ell$ used during the last customization.

This setup provides an alternative way of integrating live traffic compared to the approach described in Section~\ref{sec:live-predicted-traffic}.
Changes to the live traffic situation are handled by changing $w_\ell$ via a customization.
Additional features such as turn restrictions can be implemented by adjusting $w_q$.
For severe traffic events, the three-phase approach will be significantly faster than the one described in Section~\ref{sec:live-predicted-traffic}.
However, it comes at the expense of a more complex algorithm.
Further, integrating predicted traffic becomes difficult, as $w_q$ may not drop below the $w_\ell$ with integrated congestion.

% One could also apply live traffic updates to $w_q$ immediately and only update $w_\ell$ once the difference between $w_\ell$ and $w_q$ grows too big or when some edge weight update would set $w_q < w_\ell$.

% \subsubsection{Alternative Routes with the Penalty Method}

% Providing users with multiple alternative good routes to choose from is another relevant feature for practical routing applications.
% The penalty method is an established framework for computing such alternative routes~\cite{bdgs-argrn-11,krs-eepma-13,pz-iarp-13,kobitzsch2015alternative}.
% It computes an alternative graph instead of a set of alternative routes.
% This graph is constructed iteratively.
% In each iteration, a shortest path between $s$ and $t$ is computed.
% This path (or certain parts of it) is then added to the alternative graph depending on certain quality criteria.
% Finally, all edge weights of the path and possibly adjoined edges are penalized.
% This is repeated until the current path is more than $1 + \epsilon$ times longer than the original distance or until certain other stopping criteria are fulfilled.

% The running time of the penalty method is usually dominated by shortest path computations.
% The computations are performed on $G_{\ell}$ with a few increased edge weights.
% This setting seems like a natural fit for A* with CH-Potentials.
% While the alternative route schema outlined above is conceptually simple, it leaves open quite a few details which may have a strong impact on performance.
% In this work, we do not develop our own variant of the penalty method.
% Instead we aim to reproduce the configuration used by Kobitzsch in his dissertation~\cite{kobitzsch2015alternative}.
% To the best of our knowledge, this is the latest, most thoroughly engineered and evaluated iteration of the penalty method.
% It also achieves the qualitatively best results.
% Some details remain unclear from the description in~\cite{kobitzsch2015alternative}.
% Luckily, we could obtain the original implementation from the authors to resolve all open questions.
% As we use the same configuration, qualitative results are directly comparable.
% We can focus on comparing running times.

% To ease our exposition, we briefly reiterate the details of Kobitzsch's configuration.
% The stretch is limited to $\epsilon = 0.25$, i.e. for any path $P = (u,\dots,v)$ in the alternative graph $w(P) \leq (1+\epsilon) \cdot \dist(u,v)$.
% \todo{Das bedeutet, dass der Alternativgraph z.B. Zyklenfrei sein muss. Sobald es einen Zyklus gibt kann ich $w(P)$ beliebig groß machen.}
% Shortest path edges are penalized multiplicative with a factor of $\psi = 1.1$, i.e. when an edges was on the shortest path in $k$ iterations its weight will be $w_q = w_{\ell}\cdot\psi^k$.
% We also penalize edges incident to shortest paths with an asymmetric additive rejoin penalty.
% \todo{incident = nur incomming edges oder auch outgoing? rejoin klingt nur nach incomming, incident klingt nach beiden.}
% An edge $uv$ where $u$ is on a shortest path but $v$ is not will get its weight increased by $\psi_r \cdot \dist_{w_{\ell}}(s,t) \cdot \frac{\dist_{w_q}(s,u)}{\dist_{w_q}(s,t)}$.
% Analogously, an edge $uv$ where $v$ is on a shortest path but $u$ is not will get its weight increased by $\psi_r \cdot \dist_{w_{\ell}}(s,t) \cdot \frac{\dist_{w_q}(v,t)}{\dist_{w_q}(s,t)}$.
% \todo{Was ist $w_q$ in diesem Kontext?}
% \todo{Ich habe nicht verstanden was diese $\psi_r$ Formeln tun.}
% $\psi_r$ is set to 0.01.
% To avoid over-penalization, we count the times a \emph{node} was on a shortest path and do not apply penalties, when this number exceeds $k_{\max} := \lceil\log_{\psi}(1+\epsilon)\rceil + 2$.
% Shortest paths are split into segments not yet contained in the alternative graph.
% A segment is only added if it is long enough (at least $0.05 \cdot \dist_{w_{\ell}}(s,t)$) and does not violate the maximum stretch condition.
% We terminate the algorithm when any of the following conditions becomes true:
% \begin{itemize}
% 	\item All nodes on the current shortest path have been penalized $k_{\max}$ times.
% 	\item 10 ore more segments have been added to the alternative graph.
% 	\item The shortest path is longer than $(1+\epsilon) \cdot \dist_{w_{\ell}}(s,t)$ with respect to the original metric.
% 	\item The shortest path is longer than $(1+\epsilon) \cdot \psi + 2\psi_r \cdot \dist_{w_{\ell}}(s,t)$ with respect to the penalized metric.
% \end{itemize}
% The last two stopping criteria can both be utilized for additional pruning in the A*.
% As a final optimization, we run the forward and backward search in parallel in two threads.

% In~\cite{adgw-arrn-13} Abraham et al. describe a variation of the penalty method to directly obtain alternative routes rather than an alternative graph.
% They return the first path which shares less than 80\% with the original path and has stretch less than $1+\epsilon$.
% However, their approach uses a simpler penalization schema.
% We also implement this variant with CH-Potentials but stick to the penalization schema from Kobitzsch~\cite{kobitzsch2015alternative} and denote this approach as \emph{penalty routes}.
% This leads to some qualitative improvements compared to the results reported in~\cite{adgw-arrn-13}.

\subsubsection{Temporary Driving Bans}

Truck routing differs from car routing due to night driving bans and other restrictions.
In~\cite{kswz-erptd-p-20}, a preliminary version of CH-Potentials\footnote{
In~\cite{kswz-erptd-p-20}, CH-Potentials are used as a blackbox referring to an early ArXiv preprint~\cite{strasser2019perfect} of ours.
Our submitted paper is the finished version of the ArXiv preprint.
\cite{kswz-erptd-p-20} does not describe any of the contributions of this paper.
} is used for such a scenario.
The work considers time-dependent blocked edges and waiting at parking locations.
Further, a trade-off between arrival time and route quality is considered.

\section{Evaluation}

\label{sec:experiments}

\begin{table}
\centering
\caption{Instances used in the evaluation with preprocessing running times to construct a CH or CCH-Potential. With CCH-Potentials $w_{\ell}$ can be updated by rerunning Phase 2 again.}\label{tab:graphs}
\input{table/graphs.tex}
\end{table}

In this section, we present our experimental evaluation.
Our benchmark machine runs openSUSE Leap 15.3 (kernel 5.3.18), and has 128\,GiB of DDR4-2133 RAM and an Intel Xeon E5-1630 v3 CPUs which has four cores clocked at 3.7\,Ghz and 4~$\times$~32\,KiB of L1, 8~$\times$~256\,KiB of L2, and 10\,MiB of shared L3 cache.
All experiments were performed sequentially.
Our code is written in Rust and compiled with rustc 1.57.0-nightly in the release profile with the target-cpu=native option.
The source code of our implementation and the experimental evaluation can be found on Github\footnote{\url{https://github.com/kit-algo/ch_potentials}}.

\subsection{Inputs and Methodology}
Our main benchmark instance is a graph of the road network of Germany obtained from Open Street Map\footnote{\url{https://download.geofabrik.de/europe/germany-200101.osm.pbf}}.
To obtain the routing graph, we use the import from RoutingKit\footnote{\url{https://github.com/RoutingKit/RoutingKit}}.
The graph has 16M nodes and 35M edges.
For this instance, we have proprietary traffic data provided by Mapbox\footnote{\url{https://mapbox.com}}.
The data includes a live traffic snapshot from Friday 2019/08/02 afternoon and comes in the form of 320K OSM node pairs and live speeds for the edge between the nodes.
It also includes traffic predictions for 38\% of the edges as predicted speeds for all five minute periods over the course of a week.
We exclude speed values which are faster than the freeflow speed computed by RoutingKit.
We also perform experiments on the Europe instance provided by PTV\footnote{\url{https://ptvgroup.com}} for the 9th DIMACS implementation challenge~\cite{DemetrescuGJ09}.
Additionally, we have three graphs with proprietary traffic predictions also provided by PTV.
The PTV instances are not OSM-based.
One is an old instance of Germany with traffic predictions from 2006 for 7\% of the edges which was used for evaluation of many competing time-dependent algorithms.
The other two are newer instances of Europe with predictions for 27\% (TDEur17) and 76\% (TDEur20) of the edges.
Table~\ref{tab:graphs} contains an overview over the instances.
In this table, we further include the sequential preprocessing running time to construct the (C)CH.
We report preprocessing running times as averages over 10 runs.
To evaluate point-to-point queries, we generate 10\,000 queries where both source and target are nodes drawn uniformly at random and report average results.
For many-to-one algorithms, we use the methodology from~\cite{delling_et_al:OASIcs:2011:3266}.
We generate a \emph{ball} $B$ of nodes around 100 uniformly at random chosen nodes by running a Dijkstra search until $|B|$ nodes are settled where $|B|$ is a parameter varied throughout the experiments.
From each ball we uniformly at random pick 100 targets and a set of sources $S$ with $|S| = 2^{14}$.
We perform a many-to-one query for each target computing distances from all sources to it.
This gives us 100 many-to-one queries per ball and a total of 10\,000 one-to-many queries per ball size.

\subsection{Lazy RPHAST}\label{sec:exp_lazy_rphast}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/lazy_rphast_inc.pdf}
\caption{
Average running times of incremental Lazy RPHAST while querying $|S| = 2^{14}$ from a ball of varying size $|B|$ on OSM Germany excluding selection times.
The upper figure contains the total elapsed running time.
The lower figure contains the averaged running time per source, i.e. $y/x$ from the upper figure.
Note the different y-axis scales and units.
}\label{fig:lazy_rphast_inc}
\Description{
Two line graphs with each eight lines for different ball sizes and algorithms.
The x-axis in both graphs is the number of nodes that were queried up to that point in the incremental query.
In each graph, all lines have a similar shape and appear as scaled variants of each other.
The bigger the ball size the slower the whole algorithm.
The CCH based variant is somewhat slower for the most part.
Only for small ball sizes it is marginally faster when computing distances for more than 4096 sources.
When the ball is the entire graph, the CH based variant takes around 2.5ms for the first 512 sources, 5ms for 2048 sources and 25ms for 16384 sources.
Conversely, the average time per query is around 100mus for the first few queries, then drops to around 5mus after 512 sources, down to 2mus after 16384 sources.
For the smallest ball size of 16384, everything is 1.5 orders of magnitude faster.
}
\end{figure}

To evaluate Lazy RPHAST in the incremental setting, we measure the elapsed running time after distances from $2^i$ sources were queried.
Figure~\ref{fig:lazy_rphast_inc} depicts the total elapsed time and the average running time to compute a single distance.
The first few distances are the most expensive ones since a lot of the CH search space has not been explored yet.
With around 100\,$\mu$s, the running times are comparable to standard CH queries.
For later queries, only little work remains to be done and the overhead per distance becomes almost constant depending on the ball size.

The performance difference between the CH and CCH based variant is interesting.
The CCH based variant can utilize the elimination tree for a more efficient implementation but the search space denser.
This makes earlier queries more expensive and later queries cheaper.
With this trade-off, it depends on the ball size which variant is faster in terms of total running time.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/lazy_rphast_many_to_one_both.pdf}
\caption{
Running times of (C)CH-based Lazy RPHAST for many-to-one queries with $|S| = 2^{14}$ sources picked from a ball of varying size $|B|$.
The running time includes the selection and the time to compute all distances.
}\label{fig:many_to_one}
\Description{
A double logarithmic line plot with four lines with similar shapes and absolute values.
Running times for both algorithms and graphs with a ball size of 16384 is around 1ms.
For a ball size of 2 to the 20 CH based variants need 10ms and CCH based variants slightly more.
For a ball size of 2 to the 24 CH based variants are at 30ms and CCH based variants slightly above 40ms.
}
\end{figure}

Despite the laziness being the distinguishing feature in Lazy RPHAST, the algorithm can still be used for non-incremental many-to-one queries.
Figure~\ref{fig:many_to_one} depicts average running times to compute distances between one target node and $2^{14}$ sources for different ball sizes.
This experiment uses the exact methodology described in~\cite{delling_et_al:OASIcs:2011:3266} which allows to roughly relate our results to the performance of other one-to-many algorithms.
Keep in mind that algorithms like RPHAST optimize for a different setting than we do.
We can compare the performance for fixed $S$-$t$ terminal sets.
The difference is that with RPHAST one can efficiently compute distances from a different $t'$ to the same $S$ set while with Lazy RPHAST one can efficiently extend $S$ while $t$ stays the same.

When looking at total running times for a single many-to-one query including selection times, Lazy RPHAST delivers very competitive performance and is as fast if not faster than all algorithms evaluated in~\cite{delling_et_al:OASIcs:2011:3266}.
For example, Delling et al. report an average RPHAST running time of 1.97\,ms for selection and query combined for $|B| = 2^{14}$ and 28.52\,ms for $|B| = 2^{20}$ on DIMACS Europe.
Lazy RPHAST (on a CH) takes 0.96\,ms and 10.76\,ms respectively, to compute the same distances.
The CCH variant is slightly slower (1.00\,ms and 14.41\,ms) but still significantly faster than RPHAST.
We conclude that that Lazy RPHAST is a very useful extension to RPHAST.
It allows efficiently handling dynamic source sets and is even faster than RPHAST in the setting where both the target and the source set change between queries.
% This is of course not a fair comparison since the approaches in~\cite{delling_et_al:OASIcs:2011:3266} are geared towards having one expensive selection phase and than many faster queries.
% When computing distances from different sources towards the same target set, RPHAST will outperform Lazy RPHAST.
% However, whenever the target set may change between queries, Lazy RPHAST offers a simple and fast alternative.

% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{fig/lazy_rphast_many_to_many.pdf}
% \caption{
% Average running times of Lazy RPHAST for many-to-many queries with $|S| = |T| = 2^{14}$ sources and targets picked from a ball of varying size $|B|$ with standard deviation indicated as error bars.
% The lower part of the bars indicates the running time of the Bucket CH subphase.
% }\label{fig:many_to_many}
% \end{figure}

% Applying Lazy RPHAST to the many-to-many problem results in running times as depicted in Figure~\ref{fig:many_to_many}.
% The results are again at least competitive to the performance reported for SSE RPHAST in~\cite{delling_et_al:OASIcs:2011:3266}.
% % TODO konkrete zahlen?
% The Bucket CH based first query phase of our implementation takes between half of the total query time for $|B| = 2^{14}$ and a tenth of the time for $|B| = 2^{24}$.
% It is worth noting that contrary to SSE RPHAST, we did not explicitly utilize SIMD instructions.
% However, the code for Lazy RPHAST on an arrays of distances can be trivially vectorized by the compiler.
% We confirmed that the code was indeed vectorized by inspecting the generated assembly.

% \subsubsection{k-Nearest-Neighbor Queries with CCH}

% One-to-many queries appear in many extended applications as a subproblem.
% In some of these cases Lazy RPHAST can be used to achieve significant speed-ups with surprisingly little effort.
% One such example are k-Nearest-Neighbor Queries with CCH~\cite{buchhold_et_al:LIPIcs.SEA.2021.18}.
% There, shortest distances are computed repeatedly from the same source to boundary nodes of a cell of nodes.
% On our suggestion, the authors integrated Lazy RPHAST into their algorithm which took only a couple of lines of code\footnote{\url{https://github.com/vbuchhold/routing-framework/commit/c739dc8e81}}.
% We reran their improved experiments and include the results in our appendix.
% Briefly summarized, integrating Lazy RPHAST into this algorithm decreased running times by up to 15 times depending on the distribution of the target set for almost no effort.
% This demonstrates that Lazy RPHAST is not only useful as an A* potential but also applicable in other contexts.

\subsection{(C)CH-Potentials Heuristic}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/scaled_weights.pdf}
\caption{
Running times on a logarithmic scale for queries on OSM Ger with scaled edge weights $w_q = \alpha \cdot w_\ell$.
The boxes cover the range between the first and third quartile.
The band in the box indicates the median, the diamond the mean.
The whiskers cover 1.5 times the interquartile range.
All other running times are indicated as outliers.
}\label{fig:scaled_weights}
\Description{Fully described in the text.}
\end{figure}

The performance of A* depends on the tightness of the heuristic and the overhead of evaluating the heuristic.
CH-Potentials computes optimal distance estimates with respect to $w_\ell$.
However, for most applications, there will be a gap between $w_q$ and $w_\ell$ (otherwise one could use CH without A*).
We evaluate the impact of the difference between $w_q$ and $w_\ell$ on the performance of A*.
The lower bound $w_\ell$ is set to the freeflow travel time.
The query weights $w_q$ are set to $\alpha \cdot w_\ell$, where $\alpha\ge 1$.
% When $\alpha = 1$, the heuristic is tight.
Increasing $\alpha$ degrades the heuristic's quality.
Figure~\ref{fig:scaled_weights} depicts the results.
Clearly, $\alpha$ has significant influence on the running time.
Average running times range from below a millisecond to a few hundred milliseconds depending on $\alpha$.
Up to around $\alpha = 1.1$ the running time grows quickly.
For $\alpha > 1.1$, the growth slows down.
This illustrates both the strengths and limits of our approach and goal directed search in general.
CH-Potentials can only achieve competitive running times if the application allows for a sufficiently tight lower bounds at preprocessing time.

We observe that the running times for a fixed $\alpha$ vary strongly.
This is an interesting observation, as with uniform source and target sampling, nearly all queries are long-distance.
The query distance is thus not the reason.
After some investigation, we concluded that this is due to non-uniform road graph density.
Some regions have more roads per area than others.
The number explored A* nodes depends on the density of the search space area.
As the density varies, the running times vary.

\begin{table}
\centering
\caption{Average query running times and number of queue pushs with different heuristics and optimizations on OSM Ger with $w_q = 1.05 \cdot w_\ell$.}\label{tab:building_blocks}
\input{table/building_blocks.tex}
\end{table}

Table~\ref{tab:building_blocks} depicts the performance of unidirectional A* with different heuristics and optimizations.
We compare (C)CH-Potentials to three other heuristics.
The first heuristic is, the \emph{Zero} heuristic where $h(v)$ is $0$ for all nodes $v$.
This corresponds to using Dijkstra's algorithm.
Secondly, we compare against our own implementation of ALT~\cite{gw-cppsp-05}.
We use 16 landmarks generated with the avoid strategy~\cite{gw-cppsp-05} and activate all during every query.
Finally, we compare against a hypothetical \emph{Oracle-A*} heuristic.
This heuristic has instant access to a shortest distance array with respect to $w_\ell$, i.e. it is faster than the fastest heuristic possible in our model.
We fill this array before each query using a reverse Dijkstra search from the target node.
In this experiment, we do not account for the running time of Dijkstra's algorithm.
Thus, the reported running times of Oracle-A* do \emph{not} account for any heuristic evaluation.
(C)CH-Potentials compute the same distance estimates, but the heuristic evaluation has some overhead.
Comparing against Oracle-A* allows us to measure this overhead.
No other heuristic, which only has access to the preprocessing weights, can be faster than Oracle-A*.

We observe that the number of queue pushes roughly correlates with running time.
As (C)CH-Potentials and Oracle-A* have the same heuristic values, the number of queue pushes are by construction equal.
Each optimization reduces both queue pushes and running times.
All optimizations yield a combined speed-up of around 3.
% Skipping degree three nodes has the smallest impact.
(C)CH-Potentials outperform ALT by a factor of between six and seven and settle correspondingly fewer nodes.
This is not surprising, since ALT computes worse distance estimates.
In contrast, CH-Potentials already compute exact distances with respect to $w_\ell$.
The number of popped nodes is the same for CH-Potentials, CCH-Potentials and Oracle-A*.
The only difference between (C)CH-Potentials and Oracle-A* is the overhead of the heuristic evaluation.
This overhead leads to a slowdown of around 1.6.
Thus, CH-Potentials are already very close to the best possible heuristic in this model.
% Very interesting is the comparison with Oracle-A*.
% Our algorithm is only 1.6 times slower than accessing the exact $w_\ell$ distances instantly from a precomputed array.
This means that no competing algorithm such as ALT or CPD-Heuristics can be significantly faster.
CCH-Potentials are slightly slower than CH-Potentials because they use a weight independent node importance order.
This enables CCH-Potentials to support a customization whereas CH-Potentials cannot.

\subsection{Bidirectional A*}

\begin{table}
\centering
\caption{
Performance of different variants of bidirectional A* on OSM Ger with $w_q = 1.05 \cdot w_\ell$.
All variants alternate between the forward and the backward search.}
\label{tab:bidir_pruning}
\input{table/bidir_pruning.tex}
\end{table}

In this section, we investigate the performance of bidirectional A*.
We first evaluate different variants of bidirectional A* in Table~\ref{tab:bidir_pruning} and~\ref{tab:bidir_switching} then compare the best ones against unidirectional A* in Table~\ref{tab:bidir}.
Table~\ref{tab:bidir_pruning} depicts the performance of different variants of bidirectional A* with and without our improved pruning.
As observed in the previous section, enabling the low degree optimization achieves a speed-up by roughly a factor of three.
Symmetric bidirectional A* without our improved pruning has the worst performance.
Enabling the improved pruning significantly improves the performance of symmetric bidirectional A*.
For all heuristics except ALT, symmetric A* with the improved pruning has smaller search spaces than the average potential and very similar running times.
Without the low degree improvements, the improved symmetric variant is marginally faster and with the low degree improvements the average potential remains slightly faster.
This is due to the reduced impact of the heuristic evaluation overhead with the low degree optimizations.
Enabling the improved pruning for the average potential marginally reduces the search space size and slightly increases running times.

\begin{table}
\centering
\caption{
Performance of different direction selection criteria of bidirectional A* on OSM Ger with different query weights.
The symmetric variant uses the improved pruning, the average variant does not.
All variants use all low degree optimizations.
}\label{tab:bidir_switching}
\input{table/bidir_switching.tex}
\end{table}

Table~\ref{tab:bidir_switching} shows the performance of bidirectional A* with different strategies to decide whether to advance the forward or the backward search next.
The results clearly show that alternating the directions is always superior.
Selecting the direction by minimum queue key may lead to huge imbalances in the progress of the searches.
This causes the searches to meet later and the total search space to grow significantly.

\begin{table}
\centering
\caption{
Performance of bidirectional and unidirectional A* on OSM Ger with different query weights.
The symmetric variant uses the improved pruning, the average variant does not.
All variants use all low degree optimizations.
}\label{tab:bidir}
\input{table/bidir.tex}
\end{table}

In Table~\ref{tab:bidir} we investigate the effectiveness of bidirectional search compared to unidirectional search depending on the query weights.
Interestingly, only the zero heuristic and ALT consistently achieve speed-ups through bidirectional search on all query weights.
With $w_q = w_{\ell}$, unidirectional CH-Potentials are already optimal and only traverse the shortest path.
Here, bidirectional search will only introduce unnecessary overhead.
When query weights are scaled up uniformly, bidirectional search achieves some search space reduction, but it is not enough to significantly reduce running times due to the overhead of running a second search.
This changes drastically when the query weight increases are applied non-uniformly in the third scenario.
Here only weights with a speed less than 80\,kph were scaled up.
This touches only the beginning and end of most shortest paths between randomly chosen nodes.
The middle part of the shortest paths will typically use faster edges like highways.
In this case, bidirectional (C)CH-Potentials are a factor of two faster than the unidirectional variant.
The reason for this is that the search space of a unidirectional search expands greatly while exploring the end of the path to the target where the reduced weights are bad.
In contrast, the bidirectional searches meet in the middle of the shortest path where the reduced weights are close to zero and thus avoid this expansion.
This is also the reason why ALT behaves like this for all query weights.
By construction, the ALT heuristic has better reduced weights for edges which are on many shortest paths like highways.
Conversely, unimportant edges have bad reduced weights.
This makes bidirectional search so critical for ALTs performance.
In contrast, having a potential as tight as CH-Potentials makes bidirectional search in many scenarios unnecessary.
Bidirectional search for CH-Potentials only pays of when the reduced weights exhibit certain very specific patterns.

\subsection{Applications}

\begin{table*}
\centering
\caption{
CH-Potentials performance for different route planning applications.
Depending on the problem we apply unidirectional or bidirectional CH-Potentials (CH U or CH B) or CCH-Potentials (CCH U/B).
We report average running times and number of queue pushes.
We also report the average length increase, that is how much longer the final shortest distance is compared to the lower bound.
Finally, we report the average running time of Dijkstra's algorithm as a baseline and the speedup over this baseline.
}\label{tab:applications}
\input{table/applications.tex}
\end{table*}

Table~\ref{tab:applications} depicts the running times of (C)CH-Potentials in various applications, such as those described in Section~\ref{sec:extensions}.
We report speedups compared to extensions of Dijkstra's algorithm for each application respectively.
We start with the base case where $w_q = w_\ell$.
This is the problem variant solved by the basic CH algorithm.
CH achieves average query running times of 0.16\,ms on OSM Ger.
CH-Potentials are roughly four times slower but still achieve a huge speedup of 3795 over Dijkstra.
Such large speedups are typical for CH.
This shows that CH-Potentials gracefully converges toward a CH in the $w_q = w_\ell$ special case.
On DIMACS Europe, the average degree is somewhat higher which makes the low-degree optimizations less effective which leads to more queue operations and a slightly slower running time.

In the other scenarios, the performance of CH-Potentials strongly depends on the quality of the heuristic.
We measure this quality using the length increase of $w_q$ compared to $w_\ell$.
Forbidding highways results in the largest length increase and in the smallest speedup.
The other extreme are turn restrictions.
They have only a small impact on the length increase.
The achieved speedups are therefore comparable to CH speedups.
Since turn costs and restrictions appear mostly in the beginning and end of shortest-paths and not in the middle on highways, utilizing bidirectional search results in even better speed-ups.
Mapbox live traffic has a length increase of around 15\%, which yields running times of 130\,ms.
Applying bidirectional CH-Potentials in this case is actually detrimental to the performance because the bad reduced edge weights appear in the middle of shortest paths.
The same is the case for forbidden tunnels or highways.
Applying CCH-Potentials customized to the current traffic (or the metric with forbidden edges) to these problems yields running times similar to the unmodified case and speed-ups of more than three orders of magnitude.
With bidirectional CCH-Potentials, one can easily additionally support turn costs and still have running times of a few milliseconds without any adjustments to the customization.

The length increase of Mapbox traffic predictions are about 12.3\%, and results in a running time of 120\,ms.
The speedup in the predicted scenario is larger than in the live setting, as the travel time function evaluations slow down Dijkstra's algorithm.
Combining predicted and live traffic results in a running time of around 200\,ms.
Further adding turn restrictions, significantly increases the running times.
This increase is mostly due to the BCC optimization of Section~\ref{sec:largested-biconnected-component} becoming ineffective when considering turns.
It is not due to the length increase of using turns.
With everything activated, our algorithm still has a speedup of 13.5 over the baseline.
Interestingly, the PTV traffic predictions have a much smaller length increase than the Mapbox predictions.
This results in somewhat smaller running times of our algorithm.

% \subsubsection{Alternative Routes}

% Our implementation of the penalty method reproduces the qualitative results reported in~\cite{kobitzsch2015alternative} (for random queries on DIMACS Europe we on average achieve a total distance of 2.93, an average distance of 1.04, 5.57 decision edges and a success rate of 98.7\% with 6.7 iterations).
% The average running times to compute an alternative graph are 1.37\,s on DIMACS Europe and 0.85\,s on OSM Germany.
% These running times are roughly one order of magnitude slower than the results achieved by Kobitzsch.
% The simpler penalty routes approach is much faster and takes only 156\,ms on DIMACS Europe and 66\,ms on OSM Germany.
% This is two orders of magnitude faster than the running times reported in~\cite{adgw-arrn-13} which is not surprising since their implementation uses an unaccelerated Dijkstra.
% Due to the advanced penalization schema, our average qualitative results differ from~\cite{adgw-arrn-13}: the success rate is with 96.8\% slightly higher, the UBS with around 9\% slightly worse, the sharing is with 42\% somewhat higher but the local optimality, which was the weakest point in the original approach, with 21\% significantly better.
% Finding these alternatives takes only 1.44 iterations on average.
% This makes CH-Potential-based penalty routes a simple and feasible practical approach to compute alternative routes.

\section{Conclusion}
\label{sec:conclusion}

In this article, we introduced Lazy RPHAST, an efficient algorithm for the incremental many-to-one problem.
We also showed that it delivers competitive performance for many-to-one problems.
Based on Lazy RPHAST, we developed (C)CH-Potentials, a fast, exact, and flexible two-phase algorithm based on A* and (C)CH for finding shortest paths in road networks.
The approach can handle a multitude of complex, integrated routing scenarios with very little implementation complexity.
CH-Potentials provides \emph{exact} distances with respect to lower bound weights known at preprocessing time as an A* heuristic.
Thus, the query performance of CH-Potentials crucially depends on the availability of good lower bounds in the preprocessing phase.
Our experiments show, that this availability highly depends on the application.
We also show that the overhead of our heuristic is within a factor 1.6 of a hypothetical A*-heuristic that can instantly access lower bound distances.
Achieving significantly faster running times could still be possible in variations of the problem setting.
This leads to multiple avenues for future research.

Dropping the provable exactness requirement using a setup similar to anytime A*~\cite{DBLP:conf/aaai/ZhouH02,DBLP:conf/nips/LikhachevGT03} would be interesting.
Another promising research avenue would be to investigate graphs other than road networks.
A lot of research into grid maps exists including a series of competitions called GPPC~\cite{DBLP:conf/socs/SturtevantTTUKS15}.
Hierarchical techniques have been shown to work well on these graphs~\cite{DBLP:conf/aaai/UrasK14}.
It might also be worthwhile to apply CH-Potentials to even more routing applications, for example to the shortest $\epsilon$-smooth path problem~\cite{dss-tarrn-18} or the problem of computing alternative routes~\cite{adgw-arrn-13,bdgs-argrn-11,kobitzsch2015alternative}.
Many-to-one problems also appear as subproblems in other routing algorithms, for example in nearest neighbor computations~\cite{buchhold_et_al:LIPIcs.SEA.2021.18}.
Investigating whether Lazy RPHAST can be used to improve these algorithms appears to be a worthwhile direction for future research.
Studying the performance of Lazy RPHAST in a many-to-many context would also be interesting.

%%
%% Bibliography
%%

%% Please use bibtex,

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

% \appendix

% \section{Improved Results for k-Nearest-Neighbor Queries on CCH}

% In this section, we report the results of the k-Nearest-Neighbor Queries on CCH with Lazy RPHAST.
% The authors of~\cite{buchhold_et_al:LIPIcs.SEA.2021.18} provide the code for their algorithm and their experiments on Github\footnote{\url{https://github.com/vbuchhold/routing-framework}}.
% This includes the code for both the old version and the new one improved with Lazy RPHAST.
% We reran the experiments performed for~\cite{buchhold_et_al:LIPIcs.SEA.2021.18} in both the old and the new configuration on our machine.
% In the results, we denote the old version as \emph{CCH/CRAD} and the new one as \emph{CCH*/CRAD*}.
% Our results exhibit the same characteristics as observed in the original publication.
% Absolute running times of course vary slightly.

% Lazy RPHAST yields the greatest improvements of more than an order of magnitude when the points of interest (\emph{POIs}) are concentrated in a smaller region (see Table~\ref{tab:knn} and Figure~\ref{fig:knn_ball_size}).
% When the POIs are distributed over the whole graph, the speed-up is only around factor of three.
% Lazy RPHAST also achieves greater speed-ups for larger numbers of POIs (see Figure~\ref{fig:knn_num_pois}).
% For the demand generation application, this translates to a speed-up factor of around 2.5 (see Figure~\ref{fig:knn_demands}).
% For a more detailed discussion of these results and a comparison to competing algorithms, we refer to~\cite{buchhold_et_al:LIPIcs.SEA.2021.18}.

% \begin{table}
% \centering
% \caption{
% Performance of different closest-POI algorithms for various POI distributions.
% For each distribution, we report the time to index a set of POIs (selection time), the space consumed by the
% index (selection space), and the time to find the $k = 1, 4, 8$ closest POIs (query time).
% }\label{tab:knn}
% \input{table/knn.tex}
% \end{table}

% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{fig/knn_ball_size.pdf}
% \caption{
% Selection and query times of various closest-POI algorithms with $|P| = 2^{14}$ POIs picked at random from a ball of varying size $|B|$.
% Queries find the $k = 4$ closest POIs.
% }\label{fig:knn_ball_size}
% \end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{fig/knn_num_pois.pdf}
% \caption{
% Selection and query times of various closest-POI algorithms with a varying number $|P|$ of POIs picked at random from a ball of size $|B| = |V|$.
% Queries find the $k = 4$ closest POIs.
% }\label{fig:knn_num_pois}
% \end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=.6363\linewidth]{fig/knn_demands.pdf}
% \caption{
% Time to generate a single trip with different demand generators for various values of $\lambda$.
% }\label{fig:knn_demands}
% \end{figure}

\end{document}