### 2022-07-21 paper/table/graphs.tex

\begin{tabular}{lrrrrr}
\toprule
 &                &                & \multicolumn{3}{c}{Preprocessing [s]} \\ \cmidrule(lr){4-6} & Vertices          & Edges          & \multirow{2}{*}{CH} & \multicolumn{2}{c}{CCH} \\ \cmidrule(lr){5-6} & $[\cdot 10^6]$ & $[\cdot 10^6]$ &                     & Phase 1 & Phase 2 \\
\midrule
OSM Germany   &       16.2 &       35.4 &                          298.7 &        1\,467.4 &          10.1 \\
DIMACS Europe &       18.0 &       42.2 &                          276.2 &        2\,480.9 &          12.4 \\
TDGer06       &        4.7 &       10.8 &                           59.2 &         331.7 &           2.7 \\
TDEur17       &       25.8 &       55.5 &                          293.9 &        2\,102.3 &          14.1 \\
TDEur20       &       28.5 &       60.9 &                          311.9 &        2\,219.5 &          15.2 \\
\bottomrule
\end{tabular}